Slitasje og kjørespor

### 2022-12-19 DRAFT_slitasjeOgKjørespor.Rmd
plan

#### Infrastructure index
The infrastructure index is explained [here](https://brage.nina.no/nina-xmlui/handle/11250/2983607).
I use it as a kind of pressure or human influence index. It is likely to be a good predictor for *slitasje*.

@@

Import a stars proxy (no data imported yet)
```{r, eval=F}
infra <- stars::read_stars(paste0(path, "/NY_INFRA_IND.tif"))
```

It's easier to see what's happening if we zoom in a bit. 
Lets get a boundary box around Trondheim.
```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra))
```

Cropping the raster to speed up processing.
```{r}
infra_trd <- sf::st_crop(infra, myBB)
```

Transform to lat long dye to osm requirements
```{r, warning=F}
infra_trd_ll <- sf::st_transform(infra_trd, 4326)
```

Get the boundary box of the cropped raster
```{r}
myBB_ll <- sf::st_bbox(infra_trd_ll)
```

Download highways using the above bbox.
```{r, warning=FALSE, message=F, eval=F}
hw <- 
  osmplotr::extract_osm_objects(
    bbox = myBB_ll,
    key = "highway",
    sf = T)
```

Transforming highway data back into utm, although not strictly necessary.
```{r, eval=F}
hw_utm <- sf::st_transform(hw, sf::st_crs(infra_trd)) 
```

This object contains too many roads. I'll take out the unnamed roads. 
```{r, eval=F}
hw_utm <- hw_utm[!is.na(hw_utm$name),]
```

Caching the sf object
```{r, eval=F}
saveRDS(hw_utm, "data/cache/highways_trondheim.rds")
```

Load object
```{r}
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

@@

The infrastructure index is in EPSG32633. I will need to transform it. It takes a couple of minutes, so I will store the result in P and load it back in afterwards.
```{r, eval=F}

@@

<!-- Text and analyses here -->

I can scale the indicator for each polygon, or I can chose to aggregate
them first. If the scaled value is representative and precise at the
polygon level, then I could scale at that level. However, the combined
surveyed area is still so little of the total area of Norway, so that
only producing indicator values for these leaves the indicator without
much value for regional assessments. Later when we produce assessments
and indicator values at the regional level, we cannot simply do an area
weighting of the polygons in each region. This is because we don't want
to assume that the polygons are representative far outside of the mapped
area. But perhaps we can assume them to be representative inside
*homogenous ecological areas*. That's where the infrastructure index comes in. I will scale the indicators at the polygon level. I then take the simplified infrastructure index and, one class at the time, extract the indicator value that intersect with those cells, and extrapolate the mean of those values to all similar cell within a given region.

@@

<!-- Text here -->
Bootstrapping of nin-polygons per infrastructure class and region?

>>>>

#### Infrastructure index

The infrastructure index is explained
[here](https://brage.nina.no/nina-xmlui/handle/11250/2983607). I use it
as a kind of pressure or human influence index. It is likely to be a
good predictor for *slitasje*.

@@

Import a stars proxy (no data imported yet)

```{r, eval=F}
infra <- stars::read_stars(paste0(path, "/NY_INFRA_IND.tif"))
```

It's easier to see what's happening if we zoom in a bit. Lets get a
boundary box around Trondheim.

```{r}
myBB <- st_bbox(c(xmin=260520.12, xmax = 278587.56,
                ymin = 7032142.5, ymax = 7045245.27),
                crs = st_crs(infra))
```

Cropping the raster to speed up processing.

```{r}
infra_trd <- sf::st_crop(infra, myBB)
```

Transform to lat long dye to osm requirements

```{r, warning=F}
infra_trd_ll <- sf::st_transform(infra_trd, 4326)
```

Get the boundary box of the cropped raster

```{r}
myBB_ll <- sf::st_bbox(infra_trd_ll)
```

Download highways using the above bbox.

```{r, warning=FALSE, message=F, eval=F}
hw <- 
  osmplotr::extract_osm_objects(
    bbox = myBB_ll,
    key = "highway",
    sf = T)
```

Transforming highway data back into utm, although not strictly
necessary.

```{r, eval=F}
hw_utm <- sf::st_transform(hw, sf::st_crs(infra_trd)) 
```

This object contains too many roads. I'll take out the unnamed roads.

```{r, eval=F}
hw_utm <- hw_utm[!is.na(hw_utm$name),]
```

Caching the sf object

```{r, eval=F}
saveRDS(hw_utm, "data/cache/highways_trondheim.rds")
```

Load object

```{r}
hw_utm <- readRDS("data/cache/highways_trondheim.rds")
```

@@

The infrastructure index is in EPSG32633. I will need to transform it.
It takes a couple of minutes, so I will store the result in P and load
it back in afterwards.

```{r, eval=F}

@@

I can scale the indicator for each polygon, or I can chose to aggregate
them first. If the scaled value is representative and precise at the
polygon level, then I could scale at that level. However, the combined
surveyed area is still so little of the total area of Norway, so that
only producing indicator values for these leaves the indicator without
much value for regional assessments. Later when we produce assessments
and indicator values at the regional level, we cannot simply do an area
weighting of the polygons in each region. This is because we don't want
to assume that the polygons are representative far outside of the mapped
area. But perhaps we can assume them to be representative inside
*homogenous ecological areas*. That's where the infrastructure index
comes in. Here's the plan:

1.  scale the indicators at the polygon level.

2.  take the simplified infrastructure index and, one class at the time,
    extract the indicator value that intersect with those cells, and
    extrapolate the mean of those values to all similar cell within a
    given region.

3.  calculate an indicator value for a region by doing a weighted
    averega based o the relative area of each infrastructure class.

@@

Bootstrapping of nin-polygons per infrastructure class and region?
